---
title: "A3_Analysis_GABRIEL"
author: "Gabriel Tarriba and Natalia Alvarado"
date: "November 12, 2015"
output: html_document
---

Governance around the World
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
# Set working directory
getwd()
try(setwd("/Users/Natalia/Documents/GitHub/Third-Assignment"), silent = TRUE)
try(setwd("/Users/Gabriel/Desktop/Third-Assignment"), silent = TRUE)

# Load necessary packages, if you do not have this packages installed we highly recommend it
library(ggplot2)
library(repmis)
library(knitr)
library(plm)
library(Hmisc)
library(ggplot2)
library(RColorBrewer)
# Load Data
merged <- read.csv(file="MergedData.csv")




library(rworldmap)
world_corruption <- joinCountryData2Map(merged
                             ,joinCode = "ISO2"
                             ,nameJoinColumn = "iso2c"
                             ,mapResolution = 'coarse')
     
colorpalette <- brewer.pal(7,'RdYlBu')
world_gov <- mapCountryData(world_corruption, nameColumnToPlot='corrupest', mapTitle= 'Governance around the World',
               colourPalette = colorpalette)
print(world_gov)
```



1. Analysis of simple correlation between Control of Corruption and Natural Resource Dependence

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
# Set working directory
getwd()
try(setwd("/Users/Natalia/Documents/GitHub/Third-Assignment"), silent = TRUE)
try(setwd("/Users/Gabriel/Desktop/Third-Assignment"), silent = TRUE)

# Load necessary packages, if you do not have this packages installed we highly recommend it
library(ggplot2)
library(repmis)
library(knitr)
library(plm)
library(Hmisc)
library(ggplot2)

# Load Data
merged <- read.csv(file="MergedData.csv")

#Set data as panel data
merged <- plm.data(merged, index=c("iso2c", "year"))
```

# Central Tendency Measures

### Corruption
The mean for corruption is ***`r mean(merged$corruptest)`***

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
histcorruption <- hist(merged$corrupest,
     col = "peachpuff",
     border = "black",
     prob = TRUE,
     xlab = "Control of Corruption",
     main = 'Corruption Control Histogram')
lines(density(merged$corrupest),
      lwd = 2,
      col = "chocolate3")
abline(v = mean(merged$corrupest), 
       col = "royalblue", 
       lwd = 2)
abline(v = median(merged$corrupest),
       col = "red",
       lwd = 2)
legend(x = "topright",
	c("Density plot", "Mean", "Median"),
	col = c("chocolate3", "royalblue", "red"),
	lwd = c(2, 2, 2))
print(histcorruption)
```
### Oil Rents
The mean for the oil rents to GDP is ***`r mean(merged$oilrents)`***

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
histoilrents <- hist(merged$oilrents,
     col = "peachpuff",
     border = "black",
     prob = TRUE,
     xlab = "% of Oil Rents to GDP",
     main = 'Oil Rents to GDP Histogram')
lines(density(merged$oilrents),
      lwd = 2,
      col = "chocolate3")
abline(v = mean(merged$oilrents), 
       col = "royalblue", 
       lwd = 2)
abline(v = median(merged$oilrents),
       col = "red",
       lwd = 2)
legend(x = "topright",
	c("Density plot", "Mean", "Median"),
	col = c("chocolate3", "royalblue", "red"),
	lwd = c(2, 2, 2))
print(histoilrents)
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
histtotrents <- hist(merged$totrents,
     col = "peachpuff",
     border = "black",
     prob = TRUE,
     xlab = "% of Total Rents to GDP",
     main = 'Total Natural Resources Rents to GDP Histogram')
lines(density(merged$totrents),
      lwd = 2,
      col = "chocolate3")
abline(v = mean(merged$totrents), 
       col = "royalblue", 
       lwd = 2)
abline(v = median(merged$oilrents),
       col = "red",
       lwd = 2)
legend(x = "topright",
	c("Density plot", "Mean", "Median"),
	col = c("chocolate3", "royalblue", "red"),
	lwd = c(2, 2, 2))
print(histtotrents)
```
```{r}
# Chart 1 - Total rents vs Corruptest
ggplot(data = merged, aes(x = corrupest,
        y = totrents)) + geom_point() + theme_bw(base_size = 13)

# Chart 2 - Total rents vs Corruptest with smoothed trend line
ggplot2::ggplot(merged, aes(corrupest, totrents)) +
    geom_point() + geom_smooth() + theme_bw()

# Chart 3 - Oil rents vs Corruptest with smoothed trend line
ggplot2::ggplot(merged, aes(corrupest, oilrents)) +
    geom_point() + geom_smooth() + theme_bw()

# Examine variables
summary (merged$totrents)

# Examine Total Natural Resource Rents
summary (merged$totrents)

# Create levels of Total Natural Rent
merged$totrentgroup[merged$totrents < 1.151] <- 1
merged$totrentgroup[merged$totrents < 1.151 & merged$totrents < 5.022] <- 2
merged$totrentgroup[merged$totrents < 5.022 & merged$totrents < 12.840] <- 3
merged$totrentgroup[merged$totrents > 12.840 ] <- 4
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
# Examine GDP per capita
summary (merged$gdppc)

# Create levels of GDP per Capita
merged$gdppc[merged$gdppc
<829.2 ] <- 1
merged$gdppc[merged$gdppc
>= 829.2 & merged$gdppc < 10.740] <- 2
merged$gdppc[merged$gdppc
>= 10.740 & merged$gdppc < 13.890] <- 3
merged$gdppc[merged$gdppc
>= 13980.0] <- 4


# Try to get color chart
ggplot(data = merged, aes(corrupest, totrents)) + geom_point(aes(colour = gdppc)) +theme_bw()

```

