---
title: "Collaborative Research Project - Assignment 3"
author: "N. Alvarado and G. Tarriba"
date: "13 November 2015"
output: 
  html_document:
    fig_caption: yes
    number_sections: no
  pdf_document:
    fig_caption: yes
    number_sections: yes
  bibliography:
    - Packages.bib
    - Bibliography_A3.bib
  
runtime: shiny
---
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

# Set working directory, ensuring that code runs on either of our computers
getwd()
try(setwd("/Users/Natalia/Documents/GitHub/Third-Assignment"), silent = TRUE)
try(setwd("/Users/Gabriel/Desktop/Third-Assignment"), silent = TRUE)

# Load necessary packages, if you do not have these packages installed we highly recommend them
library(ggplot2)
library(repmis)
library(knitr)
library(plm)
library(Hmisc)
library(ggplot2)
library(texreg)
library(bibtex)

# Load Data (generated from file 'corruption_dataset.R')
merged <- read.csv(file="MergedData.csv")

# Check for nummerical variables 
str(merged)

# Set variables as numerical
merged$X <- as.numeric(merged$X)

```
##Revisiting the natural resource curse: does the relationship between natural resource rents and corruption persist into the 21st Century? 

### 1. Introduction

Our collaborative research project seeks to provide an answer to the following question: how significant is the relationship between a country’s degree of dependence on natural resource rents and its corruption level? In the 1990s and early 2000s, several scholars (see among others @leite1999, @sala2003, @pendergast2007 and @shaxson2007) identified this relationship and developed theories to account for the "Natural Resource Curse" first identified by Sachs and Werner (-@sachs1995). However, in light of the rapid economic development of many resource-dependent developing countries since the turn of the century (notably in sub-Saharan Africa, which is now the second fastest growing region in the world after Asia-Pacific @africa2014), it is worth revisiting this relationship. Perhaps it is no longer the case that the reliance on natural resource rents impairs governance and facilitates corruption. 

### 2. Data preparation

We plan to conduct analysis on data from all countries in the world for the last 20 years, because the existing empirical literature analyzes data prior to this period. As our research question suggests, we plan to use a *Corruption* indicator as our dependent variable and a measure of dependence on natural resource rents as our explanatory variable of interest. We also want to include control variables such as *GDP/GNI per capita* and *Unemployment*. These variables are included in similar empirical studies. Please see our [Assignment 2](https://github.com/natalia-alvarado/Second-Assignment) for more details on the variables to be included in our analysis, their conceptual definition and their operationalization.

We want to be create indicators to run the following model:

$$Corruption^*_{it} = \beta_1 NaturalResourceRents^*_{it} + \beta_2 GDPpercapita^*_{it} + \beta_3 Unemployment^*_{it} + \epsilon$$

### 2.1 Data sources 

The data necessary for our empirical analysis is located in two sources:

- The [World Governance Indicators (WGI)](http://data.worldbank.org/data-catalog/worldwide-governance-indicators) of the World Bank, defined as "Aggregate and individual governance indicators for 215 countries and territories over the period 1996–2014, for six dimensions of governance".

- The [World Development Indicators (WDI)](http://data.worldbank.org/data-catalog/world-development-indicators) also from the World Bank. This is "The primary World Bank collection of development indicators, compiled from officially-recognized international sources. It presents the most current and accurate global development data available, and includes national, regional and global estimates."

In [Assignment 2](https://github.com/natalia-alvara do/Second-Assignment) we also considered including in our analysis an indicator of prevalence of conflict, using data from the [UCDP/PRIO Armed Conflict Dataset](http://www.pcr.uu.se/research/ucdp/datasets/ucdp_prio_armed_conflict_dataset/). However, upon close inspection of the data we realized that it was not useful since it does not include information on the magnitude of conflicts.

### 2.2 Data gathering

- We gathered data directly from the **World Development Indicators** using the [WDI Package for R](https://cran.r-project.org/web/packages/WDI/WDI.pdf).This package enables us to automatically download WDI indicators using the World Bank's APIs.

```{r, echo=FALSE}

```

- For the **World Governance Indicators** there was no API. Hence, the data was automatically downloaded as a .XLSX file and then imported into R.

```{r, echo=FALSE}

```

## 2.3 Descriptive Statistics

```{r, echo=FALSE}

```




## 2.4 Inferencial Statistics

### Corruption over time
```{R, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
corrtime <- ggplot(aes(x = year, y = corrupest), data = merged) + geom_point(position = "jitter") + theme_bw() + geom_smooth()
```

### Regressions

Following the model

$$Corruption_i=\beta_0+\beta_1 NaturalResourceRents_i +\beta_2 GDPpercapita_i + \beta_3 Unemployment_i + \epsilon_i$$

The model one is a regression with the control variables specified.

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
model1 <- lm(corrupest ~ totrents + gdppc + unemp, data = merged)
confint(model1)
summary(model1)

```

### Heteroskedasticity
When testing for heterosckedasticity, ...

```{}r
plot(model1, which = 1, main = "Residual vs. Fitted Plot")
```
The graphical inspection suggests there is heterosckedastocity. For this reason, it is necessary


### Pearson's correlation
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}


```




$$Corruption_i=\beta_0+\beta_1 NaturalResourceRents_i +\beta_2 GDPpercapita_i + \beta_3 ArmedConflict_i + \beta_4 Unemployment_i + \epsilon_i$$

The next step will be to run a Fixed Effects regression, which will allow us to get rid of the country-specific fixed effects (such that we will only be analyzing the variation in variables within one country across time). To run a Fixed Effects regression, we will subtract from each term its mean. Thus, the regression equation will be:
$$Corruption^*_{it} = \beta_1 NaturalResourceRents^*_{it} + \beta_2 GDPpercapita^*_{it} + \beta_3 ArmedConflict^*_{it} + \beta_4 Unemployment^*_{it} + \epsilon$$

The terms in the equation above are demeaned to eliminate the fixed effects, such that:
$$NaturalResourceRents^*_{it} = NaturalResourceRents_it - \overline {NaturalResourceRents_i}$$
$$GDPpercapita^*_{it} = GDPpercapita_{it} - \overline {GDPpercapita_i}$$
$$ArmedConflict^*_{it} = ArmedConflict_{it} - \overline {ArmedConflict_i}$$
$$Unemployment^*_{it} = Unemployment_{it} - \overline {Unemployment_i}$$



```{r, echo=FALSE}
data("Wages")
?Wages
summary(Wages)
M1 <- lm(lwage ~ ed, data = Wages)
M2 <- lm(lwage ~ ed + exp, data = Wages)
summary(M1)
confint(M1)

labels <- c('')
stargazer::stargazer(M1, M2, title='OLS regression of the Percentage of Wages Variation', digits=2, type='html', header = FALSE)
type='html'
```


```{r}
# Create list of packages
PackagesUsed <- c("knitr", "ggplot2", "repmis")
# Load PackagesUsed and create .bib BibTeX file
# Note must have repmis package installed.
repmis::LoadandCite(PackagesUsed, file = "Packages.bib", install = FALSE)

```

